\documentclass[12pt]{ucsddissertation}
% mathptmx is a Times Roman look-alike (don't use the times package)
% It isn't clear if Times is required. The OGS manual lists several
% "standard fonts" but never says they need to be used.

\usepackage{tikz}
\usepackage{amsmath}

\usepackage{filecontents}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{thmtools}
\usepackage{color}
\usepackage{balance}
\usepackage{enumitem}
\usepackage{bm}
\usepackage{comment}
\usepackage{subcaption}

\usepackage[linesnumbered, lined, boxed]{algorithm2e}

\usepackage[normalem]{ulem}


\usepackage{mathptmx}
\usepackage[NoDate]{currvita}
\usepackage{array}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{ragged2e}
\usepackage{microtype}
\usepackage[breaklinks=true,pdfborder={0 0 0}]{hyperref}
\usepackage{graphicx}
\AtBeginDocument{%
	\settowidth\cvlabelwidth{\cvlabelfont 0000--0000}%
}

% OGS recommends increasing the margins slightly.
\increasemargins{.1in}

% These are just for testing/examples, delete them
\usepackage{trace}
%\usepackage{showframe} % This package was just to see page margins
\usepackage[english]{babel}
\usepackage{blindtext}
\overfullrule5pt
% ---

% Required information
\title{Private Graph Statistics and Algorithms in Modern Applications}
\author{Jacob John Imola}
\degree{Computer Science}{Doctor of Philosophy}
% Each member of the committee should be listed as Professor Foo Bar.
% If Professor is not the correct title for one, then titles should be
% omitted entirely.
\chair{Professor Kamalika Chaudhuri}
% Your committee members (other than the chairs) must be in alphabetical order
\committee{Professor Sanjoy Dasgupta}
\committee{Professor Russell Impagliazzo}
\committee{Professor Molly Roberts}
\degreeyear{2023}

% Start the document
\begin{document}
% Begin with frontmatter and so forth
\frontmatter
\maketitle
\makecopyright
\makesignature
% Optional
\begin{dedication}
\setsinglespacing
\raggedright % It would be better to use \RaggedRight from ragged2e
\parindent0pt\parskip\baselineskip
This work is dedicated to my family. First, to my parents Martha and Mario,
who inspired me to strive to be my best from a young age. Second, to my
sister Lauren and brother David, for helping and supporting me along the way. 
I love you all!
\end{dedication}
% Next comes the table of contents, list of figures, list of tables,
% etc. If you have code listings, you can use \listoflistings (or
% \lstlistoflistings) to have it be produced here as well. Same with
% \listofalgorithms.
\tableofcontents
\listoffigures
\listoftables

% Your fancy acks here. Keep in mind you need to ack each paper you
% use. See the examples here. In addition, each chapter ack needs to
% be repeated at the end of the relevant chapter.
\begin{acknowledgements}
First and foremost, I extend my sincerest gratitude to my advisor,
Professor Kamalika Chaudhuri, for making this entire journey possible.
I didn't have much direction or confidence when I arrived at UCSD,
but Kamalika saw potential in me and guided me into research. When 
something finally stuck, Kamalika let me fly more freely and allowed me
to develop my own style. She has also provided invaluable connections and
collaborations throughout the years, without which many of my research 
projects would not have been possible.

Second, I greatly appreciate my office mates 
in lab~4142; with whom I have laughed, learned, opened up, and shared
so many meaningful experiences. My PhD certainly wouldn't have been possible without 
this group. In particular, I would like to extend thanks to
Mary Anne Smart, who has helped me with many problems related to work and to 
life, and who I can always count on to be a voice of reason. Next,
thanks to Casey Meehan for his light-heartedness and generosity,
especially during the more challenging times. Next, 
thanks to Robi Bhattacharjee from whose uncanny intelligence and varied interests
I have learned lots. Finally, thanks to Amrita Roy Chowdhury, whose
extreme aptitude in research and other intellectual pursuits I find quite
inspiring and have taught me lots.

Third, I am very grateful to my collaborators over the years. My collaboration
with Takao Murakami has been my most successful one, in no small part due to 
his amazing work ethic and keen eye for interesting research projects. It has
been an absolute pleasure to work with him. Thanks to Abhinav Aggarwal, my
collaborator and manager during my time at Amazon, for setting up an interesting,
enriching research project during my time there. Thanks to Alessandro Epasto, my 
collaborator/manager at Google, who also set up an amazing research project during my
time at Google. Finally, I have benefitted greatly from working with 
Amrita Roy Chowdhury, who has an incredible amount of knowledge and research 
expertise.

Lastly, I am so grateful to my friends and family. To my friends 
from both UCSD and from Carnegie
Mellon: we have kept each other in our lives for a reason, and I hope
to share many more fun times with you all in the future. To my family: thanks
for loving and supporting me throughout my life.


Chapter 1, in full, is a reprint of the material as it appears in Jacob Imola,
Takao Murakami, and Kamalika Chaudhuri, USENIX Security Symposium, 2021.
"Locally Differentially Private Analysis of Graph Statistics." 
The dissertation author was a joint first author of this paper.

Chapter 2, in full, is a reprint of the material as it appears in
Jacob Imola,
Takao Murakami, and Kamalika Chaudhuri, 31st USENIX Security Symposium, 2022.
"Communication-Efficient Triangle Counting under Local Differential Privacy." 
The dissertation author was a joint first author of this paper.

Chapter 3, in full, is a reprint of the material as it appears in Jacob Imola,
Takao Murakami, and Kamalika Chaudhuri, Proceedings of the 2022 ACM SIGSAC Conference on Computer and Communications Security, 2022. "Differentially Private Triangle and 4-Cycle Counting in the Shuffle Model." 
The dissertation author was a joint first author of this paper.

Chapter 4, in full, is currently being prepared for submission for publication of the material.
Jacob Imola, Amrita Roy Chowdhury, and Kamalika Chaudhuri.
"Robustness of Locally Differentially Private Graph Analysis Under Data Poisoning Attacks."
The dissertation author was the first author of this paper.

Chapter 5, in full, is currently being prepared for submission for publication of the material.
Jacob Imola, Alessandro Epasto, Mohammad Mahdian, Vincent Cohen-Addad, and Vahab Mirrokni.
"Differentially Private Hierarchical Clustering with Provable Approximation Guarantees."
The dissertation author was the first author of this paper.

\end{acknowledgements}

% Stupid vita goes next
\begin{vita}
\noindent
\begin{cv}{}
\begin{cvlist}{}
\item[2018] Bachelor of Arts, Carnegie Mellon University
\item[2018-2023] Graduate Research Assistant, University of California, San
Diego
\item[2021] Research Intern, Amazon Arlington, VA.
\item[2022] Research Intern, Google N.Y.
\item[2023] Doctor of Philosophy, University of California, San Diego
\end{cvlist}
\end{cv}

% This puts in the PUBLICATIONS header. Note that it appears inside
% the vita environment. It is optional.
\publications
\begin{itemize}
\item
		Jacob Imola, Amrita Roy Chowdhury, and Kamalika Chaudhuri.
		\textit{\href{https://arxiv.org/pdf/2210.14376.pdf}{Robustness of Locally Differentially Private Graph Analysis Against Poisoning}}. Preprint, 2023.

\item
	Jacob Imola, Alessandro Epasto, Mohammad Mahdian, Vincent Cohen-Addad, and Vahab Mirrokni.
	\textit{\href{https://arxiv.org/abs/2302.00037}{Differentially-Private Hierarchical Clustering with Provable Approximation Guarantees}}. In ICML 2023.

\item
		Robi Bhattacharjee, Jacob Imola, Michal Moshkovitz, and Sanjoy Dasgupta.
		\textit{\href{https://arxiv.org/abs/2102.09101}{Online $k$-means Clustering on Arbitrary Data Streams}}. In ALT 2023.

\item
 Jacob Imola, Takao Murakami, and Kamalika Chaudhuri.
		\textit{\href{https://arxiv.org/abs/2205.01429.pdf}{Differentially Private Triangle and 4-Cycle Counting in the Shuffle Model}}. In CCS 2022.

\item Jacob Imola, Shiva Kasiviswanathan, Stephen White, Abhinav Aggarwal, Nathanael Teissier.
		\textit{\href{https://openreview.net/pdf?id=B0l8-wLjql5}{Balancing Utility and Scalability in Metric Differential Privacy}}. In UAI 2022.

\item Jacob Imola, Takao Murakami, and Kamalika Chaudhuri.
		\textit{\href{https://arxiv.org/abs/2110.06485}{Communication-Efficient Triangle Counting under Local Differential Privacy}}. In USENIX Security 2022.

\item Jacob Imola and Kamalika Chaudhuri.
		\textit{\href{https://arxiv.org/abs/2105.10594}{Privacy Amplification Via Bernoulli Sampling}}. In TPDP Workshop at ICML 2021.

\item Jacob Imola, Takao Murakami, and Kamalika Chaudhuri.
		\textit{\href{https://arxiv.org/abs/2010.08688}{Locally Differentially Private Analysis of Graph Statistics}}. In USENIX Security 2021.

\item Kamalika Chaudhuri, Jacob Imola, and Ashwin Machanavajjhala.
		\textit{\href{https://arxiv.org/abs/1907.02159}{Capacity Bounded Differential Privacy}}.
	  In NeurIPS 2019.
\end{itemize}

% This puts in the FIELDS OF STUDY. Also inside vita and also
% optional.
\end{vita}

% Put your maximum 350 word abstract here.
\begin{dissertationabstract}
In many modern machine learning applications, users hold data on a local device 
and communicate with an untrusted central analyst. Local differential privacy 
(local DP) is the state-of-the art way to ensure that all data sent by a user 
will be protected regardless of who eventually sees the data. In this dissertation,
I will focus on collecting certain graph statistics, such as the degree vector 
and number of triangles in the graph, under local DP. I will 
obtain formal performance guarantees in the presence of different real-world constraints.

First, I will detail a one-round algorithm which is simple, but not guaranteed to work 
well on sparse graphs. To improve these matters,
I will propose a two-round algorithm with improved performance 
for sparse graphs. Unfortunately, it may have prohibitively large per-user 
download cost. Second, I will show how one can use 
edge sampling to reduce download cost, and exhibit another two-round algorithm which
trades off between the download cost and error.

Third, given that it requires significant synchronization overhead to implement multiple rounds, I will 
focus on improved one-round algorithms. I propose such an algorithm which satisfies a slightly weaker
notion of DP, the shuffled model. This algorithm uses a novel wedge-shuffling approach
to count wedges, and gives rise to an accurate algorithm for counting 
4-cycles as well. Fourth, I will shift gears to computing the degree histogram in the presence
of malicious users who do not follow the local DP protocol. Many ubiquitous local DP mechanisms are susceptible to such users.
 I will detail robust algortihms which
makes use of the fact that every edge in a graph is shared by two users in order to design 
edge consistency checks which can flag malicious users.

Fifth, I will detail my work on private hierarchical clustering,
a common clustering algorithm used for graphs and other data. I propose two algorithms and 
a lower bound in the general setting of the problem, and a further algorithm in the stochastic 
block model of graphs. The algorithm design and lower bound 
are of interest to designing future local DP protocols on graphs.

\end{dissertationabstract}

% This is where the main body of your dissertation goes!
\mainmatter

% Optional Introduction
\begin{dissertationintroduction}
This optional section is barely described in the OGS manual other than
saying it is optional and that it appears in the table of contents
between the Abstract and the first chapter.

No formatting guidelines appear so presumably, it should be formatted
like an ordinary chapter. It should appear after the
\verb!\mainmatter! macro because it should start on page~1.
\end{dissertationintroduction}

% Chapter 1
\include{chapters/chapter1/chapter1}
\include{chapters/chapter2/chapter2}
\include{chapters/chapter3/chapter3}
\include{chapters/chapter4/chapter4}
\include{chapters/chapter5/chapter5}

\appendix
\include{chapters/chapter1/appendix}
\include{chapters/chapter2/appendix}
\include{chapters/chapter3/appendix}
\include{chapters/chapter4/Appendix}
\include{chapters/chapter5/7_appendix}

% Stuff at the end of the dissertation goes in the back matter
\backmatter
\bibliographystyle{plain} % Or whatever style you want like plainnat
\bibliography{references}

\end{document}
